Efficiently Stable Presentations from Error-Correcting Codes

### 2025-12-10 daj-final/stability-daj.tex
update

\documentclass[11pt]{article}
\usepackage{booktabs}
\usepackage{fullpage}
\usepackage{titlesec}
%\newcommand{\sectionbreak}{\clearpage}
\usepackage{amsmath,amsfonts,amsthm,mathrsfs,xspace,graphicx}
\usepackage[backref,colorlinks,citecolor=blue,bookmarks=true]{hyperref}
\usepackage{mathpazo}
\usepackage{nicefrac}
\usepackage{endnotes}
\usepackage{color}
\usepackage{float}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage{bbm}
\usepackage{suffix} % for *-version commands
\usepackage{times}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{amssymb,latexsym}
%\usepackage{IEEEtrantools}
\usepackage[capitalize]{cleveref}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{multirow}
\usepackage[section]{placeins}
\usepackage[affil-it]{authblk}
\usepackage{comment}
\usepackage{footmisc}

@@

% \usepackage{showlabels}
% \renewcommand{\showlabelfont}{\tiny\ttfamily\color{red}}

\bibliographystyle{alpha}

\newif\ifnotes\notestrue
%\newif\ifnotes\notesfalse

\input{marginnotes}
\begin{document}

\title{Efficiently stable presentations from error-correcting codes}

\author[1]{Michael Chapman}
\author[2]{Thomas Vidick}
\author[3]{Henry Yuen}
\affil[1]{Courant Institute of Mathematical Sciences, 
	New York University, USA}
\affil[2]{ Faculty of Mathematics and Computer Science, Weizmann Institute of Science, Israel}
\affil[3]{Department of Computer Science, Columbia University, USA}

\date{\today}
\maketitle

\noteswarning

\begin{abstract}
We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

With this goal in mind we analyze various natural examples of presentations. We provide a general method for constructing presentations of $\Z_2^k$ from linear error-correcting codes. We observe that the resulting presentation has a weak form of stability exactly when the code is  \emph{testable}. This raises the question of whether testable codes give rise to genuinely stable presentations using this method.  While we cannot show that this is the case in general, we leverage recent results in the study of non-local games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application, we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive  the quantum low-degree test of Natarajan and Vidick (IEEE FOCS'18), which is a key building block in the recent refutation of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383). 
\end{abstract}

@@

\begin{proof}
By definition, there exists a projection $P \in \mM_\infty$ and a partial isometry $w \in P \mM_\infty I_\mM$ such that $\mN = P \mM_\infty P$ and
\begin{align}
\omega(G;\strategy') &= \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( Q^x_a \, Q^y_b \big)\notag\\
&=  \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a w\, w^* Q^y_b w\big)\notag\\
&\qquad\qquad+  \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big) \notag\\
&\qquad\qquad+ \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big(Q^x_a  Q^y_b (P-ww^*)\big)~.\label{eq:cv-0}
\end{align}
Here we used the fact that $Q^x_a \in \mN$ so $P Q^x_a P = Q^x_a$ for all $x,a$. 
The second term on the right-hand side can be bounded as
%\begin{align*}
%\Big|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big)\Big|
%&=\Big| \tau\Big(  (P - w\, w^* )\Big( \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big)\Big)\Big|\\
%&\leq \tau(P-ww^*) \Big\|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big\|\\
%&\leq \eps\;,
%\end{align*}
%where the first inequality is by H\"older's inequality and the last inequality uses the definition of closeness to bound the first term. To bound the second term on the right-hand side, we write it as $\|A\circ B\|$, with $\circ$ the Hadamard product, and use  $\|A\circ B\| \leq \max_{i,j} |A_{i,j}|\|B\|$. Here $A = \sum_{a,b} D(a,b) P_{a,b}$, with $P_{a,b} = \sum_{i,j} \ket{u_{i,a}} \bra{v_{j,b}}$ with $\ket{u_{i,a}}$ a basis for the range of $Q^x_a$ and $\ket{v_{j,b}}$ for the range of $Q^y_b$, and $B=\sum_{a,b} Q^y_b ww^* Q^x_a = ww^*$.  
\begin{align}
\Big|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) & \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big)\Big|\notag\\
&\leq \Es{(x,y)\sim\mu} \Big | \tau\big( \sum_{a,b} D(a,b,x,y)   Q^y_b w w^* Q^x_a (P - w\, w^* ) \big) \Big | \notag\\
&\leq \sqrt{ \tau((P - ww^*)^2)} \cdot \sqrt{\Es{(x,y)\sim\mu} \tau\big( \big| \sum_{a,b} D(a,b,x,y)   Q^x_a w w^* Q^y_b \big|^2 \big) } \notag\\
&\leq \sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu} \tau\big( \sum_{a,b} D(a,b,x,y)   Q^y_b ww^* Q^x_a w w^* Q^y_b \big) } \notag\\
&\leq \sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu}  \sum_{a,b}  \tau\big(Q^y_b ww^* Q^x_a w w^* Q^y_b \big) } \notag \\
&=\sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu}  \sum_{b}  \tau\big(Q^y_b ww^* \big) } \notag\\
&\leq \sqrt{\eps}\;.\label{eq:cv-1}
%&=\Big| \tau\Big(  (P - w\, w^* )\Big( \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big)\Big)\Big|\\
%&\leq \tau(P-ww^*) \Big\|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big\|\\
%&\leq \eps\;,
\end{align}
The third line is due to Cauchy-Schwarz and Jensen's inequality, where $|A|^2=A^*A$. The fourth line uses that $P - ww^*$ is a positive operator with operator norm at most $1$ for the first term, and that for fixed $x,y$, the measurements $\{ Q^x_a \}_a$ and $\{Q^y_b\}_b$ are projective. The fifth line is due to $D(x,y,a,b) \in \{0,1\}$. The sixth line is due to $\sum_a Q^x_a = I_\mN$. The seventh line is due to $\sum_a Q^y_b = I_\mN$. The third time on the right-hand side of~\eqref{eq:cv-0} is bounded in the same manner.

@@

In the next section we detail an important application of this test, Proposition~\ref{prop:dim-test}, which is used in the work~\cite{ji2020mip}. The present proof somewhat simplifies the one given in~\cite{ji2020mip}, but it keeps the same overall strategy. \tnote{added for revision 2:}Moreover, Corollary~\ref{cor:qld} below is qualitatively stronger than~\cite[Theorem 7.14]{ji2020mip} as the latter applies to a slightly different game which in particular has a pair of questions with long answers, of $O(2^m\log q)$ bits, whereas the tester in Figure~\ref{fig:bRM-tester}, which is analyzed in the corollary, has short answers, $O(d)$ bits long; with a typical parameter setting (see the discussion following Theorem~\ref{thm:z2-stab}) this is exponentially smaller. The presence of long answers makes the analysis easier.

@@

To that end, assume that every genuine representation of $\Gamma$ is $(\delta,\mu_S)$-far from $\rho$. Namely, for every genuine representation $\varphi\colon \Gamma\to \mU(\mN)$, where $\mN=P\mM_\infty P$, and every isometry $w\in P\mM_\infty \Id_\mM$, we have 
\[
\max\left\{\Expectation_{s\sim \mu_S} \|\rho(s)-w^*\varphi(s)w\|_\tau^2\ ,\ \tau^\mM(\Id_\mM-w^*w)\right\}\geq \delta.
\]
Let $\varphi'\colon \Gamma\to \mU(\mN)$ be a genuine representation of $\Gamma$ which is $(\delta',\mu_S)$-close to $\rho'$, i.e.\ it satisfies 
\[
\frac{1}{|\Phi|}\Expectation_{s\sim \mu_S} \|\rho'(s)-w^*\varphi'(s)w\|_\tau^2\ ,\ \tau^{\mM_\Phi}(\Id_{\mM_\Phi}-w^*w)\leq  \delta',
\]
where  again  $\|X\|_\tau^2=\tau_\infty(X^*X)$.
As before, we can assume $w=PI_{\mM_\Phi}$. Then,
\[
\begin{split}
    \Expectation_{\alpha\in \Phi}\Expectation_{s\sim \mu_S} \|\rho(\alpha(s))-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2&=\frac{1}{|\Phi|}\Expectation_{s\sim \mu_S}\sum_{\alpha\in \Phi}\|\Id_{\mM_\alpha}\rho'(s)\Id_{\mM_\alpha}-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2 \\
    &\leq \frac{1}{|\Phi|}\Expectation_{s\sim \mu_S}\|\rho'(s)-\Id_{\mM_\Phi}\varphi'(s)\Id_{\mM_\Phi}\|_\tau^2\leq \delta'.
\end{split}
\]
In particular, by Markov's inequality, for at least two thirds of the $\alpha\in \Phi$ we have 
\[
\Expectation_{s\sim \mu_S} \|\rho(\alpha(s))-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2\leq 3\delta'.
\]
Similarly,
\[
\begin{split}
    \delta'&\geq \tau^{\mM_\Phi}(\Id_{\mM_\Phi}-\Id_{\mM_\Phi}P\Id_{\mM_\Phi})\\
    &=\Expectation_{\alpha\in\Phi}\tau_\infty(\Id_{\mM_\alpha}-\Id_{\mM_\alpha}P\Id_{\mM_\alpha})
\end{split}\;,
\]
and for at least two thirds of the $\alpha\in \Phi$ we have 
\[
\tau^{\mM_\alpha}(\Id_{\mM_\alpha}-\Id_{\mM_\alpha}P\Id_{\mM_\alpha})\leq 3\delta'.
\]
Hence, $\delta'\geq \frac{\delta}{3}$, and we deduce that every genuine representation of $\Gamma\cong \langle S\colon R\rangle$ is at least $(\nicefrac{\delta}{3},\infty)$-away from $\rho'$.
%Since the $L^\infty$ distance between functions is larger than the $L^1$ distance between them (with respect to any fully supported measure $\mu_S$), we conclude that a lower bound of $f(\eps)$ on the modulus of stability of a sufficiently symmetric presentation induces a lower bound of $\frac{\min{w_i}}{3}f(\eps)$ 
\begin{corollary}
    The transformation $\rho\mapsto\rho'$ we described  translates a $(\eps,\mu_R)$-approximate representation into a $((\max{\nicefrac{1}{w_i}})\cdot\eps,\infty)$-approximate representaion, and if every genuine representation of $\Gamma$ is $(\delta,\mu_S)$-far from $\rho$, then every genuine representation is $(\nicefrac{\delta}{3},\infty)$-far from $\rho'$.
\end{corollary}
\begin{remark}
    By applying this corollary to the construction from the beginning of the appendix, we conclude that the $L^\infty$ modulus of stability of the presentation \eqref{eq:z2-efficient} is $\Omega(k\eps)$.
\end{remark}
\bibliography{qld}

\notesendofpaper

@@

>>>>

%% daj-template.tex v0.33     23 Sep 2016   Alex Russell/Laszlo Babai
%%
%% AUTHOR: Fill in fields (or see warnings) below marked with "AUTHOR"
%% ** Add as few macro / package definitions as possible
%% ** Compile with "pdflatex"; make sure that
%%           daj.cls and tocbase.cls are in the same directory.
%%
%% EDITOR: Fill in fields below marked with "EDITOR"
%%    and check that authors proprely filled in field marked with "AUTHOR"

\documentclass{daj}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage[backref,colorlinks,citecolor=blue,bookmarks=true]{hyperref}

\usepackage{booktabs}
%\usepackage{fullpage}
\usepackage{titlesec}
%\newcommand{\sectionbreak}{\clearpage}
\usepackage{amsmath,amsfonts,amsthm,mathrsfs,xspace,graphicx}
%\usepackage{mathpazo}
\usepackage{nicefrac}
%\usepackage{endnotes}
\usepackage{color}
\usepackage{float}
\usepackage{xcolor}
\usepackage{mdframed}
\usepackage{bbm}
\usepackage{suffix} % for *-version commands
\usepackage{times}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{amssymb,latexsym}
%\usepackage{IEEEtrantools}
\usepackage[capitalize]{cleveref}
\usepackage{enumitem}
%\usepackage{tikz}
%\usepackage{tikz-cd}
\usepackage{multirow}
\usepackage[section]{placeins}
%\usepackage[affil-it]{authblk}
\usepackage{comment}
\usepackage{footmisc}

@@

\newcommand{\tnote}[1]{}
\bibliographystyle{alpha}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% AUTHOR: Fill in meta-data below:
\dajAUTHORdetails{%
  title = {Efficiently Stable Presentations from Error-Correcting Codes}, %% please capitalize all significant words
  author = {Michael Chapman, Thomas Vidick and Henry Yuen},
    %% Please use the format for commas as follows:
    %% "A", or "A and B", or "A, B, and C", or "A, B, C, and D", etc.
  plaintextauthor = {Michael Chapman, Thomas Vidick and Henry Yuen},
    %% An author list in plain text: Use the format
    %% "A", or "A, B", or "A, B, C", etc.
    %% NOTE: No LaTeX code in author names.
    %% NOTE: No "and" at the end--simply comma separated,
    % 
 %% The remaing lines in this section are optional:
    %
    %% IF YOUR TITLE CONTAINS MATH OR LATEX such as accented characters: 
    %% Add a "plain text title";  otherwise comment out the next line:
  plaintexttitle = {Efficiently Stable Presentations from Error-Correcting Codes}, %%  title without math or LaTeX
    %
    %% ONLY IF YOUR TITLE IS TOO LONG to fit in the page headers, please 
    %% add an abbreviated version of the title, otherwise comment it out:
  %runningtitle = {R\"odl's $n^{\log\log n}$ Bound}, 
    %
    %% ONLY IF YOUR AUTHOR LIST IS TOO LONG to fit in the page headers, 
    %% add an abbreviated version, otherwise comment it out:
 % runningauthor = {Paul Erd\H{o}s, Johan H{\aa}stad, L\'aszl\'o Lov\'asz, and Andrew C-C. Yao},
    %% you can replace first names and/or middle names with initials.
    %
    %% ONLY IF YOUR AUTHOR LIST IS TOO LONG to fit the copyright entry
    %% on the bottom of the front page,
    %% add an abbreviated version, otherwise comment it out:
  copyrightauthor = {M. Chapman, T. Vidick and H. Yuen},
    %% Note that the copyrightauthor  field will seldom be necessary;
    %% for instance, in this example with four authors, it would be 
    %% all right to comment it out and have all authors' full names 
    %% appear on the Copyright line
   %
   %% Include keywords of your choice: comma separated, lower case;
   %% comment out the "keywords" line if you don't wish to provide them
  keywords = {group stability, error-correcting codes},
}   %%% END \dajAUTHORdetails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% EDITOR: please fill in the following data:
\dajEDITORdetails{%
   year={20XX},
   volume={XX},
   number={XX},
   received={XX Month 20XX},   % received date: example: 7 January 2017
   revised={XX Month 20XX},    % Optional revised date (you may comment it out)
   published={XX Month 20XX},  % published date
   doi={10.19086/daXXX},       % XXX = number of paper, e.g. da006 for paper#6
%                              % or  da0006 (length of string arbitrary)
}   %%% END \dajEDITORdetails

\begin{document}

\begin{frontmatter}[classification=text]
%% EDITOR: this will force the keywords to appear right after the Abstract.
%%   If the abstract is too long and would force the keywords off the
%%   front page, please comment out % [classification=text] above
%%   This way the keywords will be floated on the bottom of the first page
%%   even though the Abstract spills over to the next page.

%%% AUTHOR: Title goes here.  This line is optional.  You must use it
%%   if title has footnote attached or requires nontrivial typesetting,
%%   e.g., inclusion of linebreaks to force nice layout.
\title{Efficiently Stable Presentations from Error-Correcting Codes} %% please capitalize all significant words

%%% AUTHOR:
%%% List all authors. If you wish, place grant acknowledgements in \thanks.
%%% In brackets include a short tag for each author.
\author[mc]{Michael Chapman\thanks{Supported by the National Science Foundation under Grant No. DMS-2424441 and Simons Foundation grant N. 965535}}
\author[tv]{Thomas Vidick\thanks{Supported by the Swiss State Secretariat for Education, Research and Innovation (SERI). Parts of this work was completed while at the Weizmann Institute, supported by a research grant from the Center for New Scientists at the Weizmann Institute of Science, AFOSR Grant No. FA9550-22-1-0391, and ERC Consolidator Grant VerNisQDevS (101086733).}}
\author[hy]{Henry Yuen\thanks{Supported by AFOSR award FA9550-23-1-0363, CCF-2144219, and the Sloan Foundation.}}

%%% AUTHOR: Abstract goes here
\begin{abstract}
We introduce a notion of \emph{efficient stability} for finite presentations of groups. Informally, a finite presentation using generators $S$ and relations $R$ is \emph{stable} if any map from $S$ to unitaries 
that approximately satisfies the relations (in the tracial norm) is close to the restriction of a representation of $G$ to the subset $S$. This notion and variants thereof have been extensively studied in recent years, in part motivated by connections to property testing in computer science. The novelty in our work is the focus on \emph{efficiency}, which, informally, places an onus on small presentations --- in the sense of encoding length.  
%(of size, say, logarithmic in the group size)
The goal in this setup is to achieve non-trivial tradeoffs between the presentation length and its modulus of stability.
%---as, intuitively, more generators and relations place more constraints on the approximate homomorphism, making stability easier to achieve---whereas we wish to minimize the number of variables and constraints.

With this goal in mind we analyze various natural examples of presentations. We provide a general method for constructing presentations of $\Z_2^k$ from linear error-correcting codes. We observe that the resulting presentation has a weak form of stability exactly when the code is  \emph{testable}. This raises the question of whether testable codes give rise to genuinely stable presentations using this method.  While we cannot show that this is the case in general, we leverage recent results in the study of non-local games in quantum information theory (Ji et al., Discrete Analysis 2021) to show that a specific instantiation of our construction, based on the Reed-Muller family of codes, leads to a stable presentation of $\Z_2^k$ of size $\poly\log(k)$ only. As an application, we combine this result with recent work of de la Salle (arXiv:2204.07084) to re-derive  the quantum low-degree test of Natarajan and Vidick (IEEE FOCS'18), which is a key building block in the recent refutation of Connes' Embedding Problem via complexity theory (Ji et al., arXiv:2001.04383). 
\end{abstract}
\end{frontmatter}

@@

\begin{proof}
By definition, there exists a projection $P \in \mM_\infty$ and a partial isometry $w \in P \mM_\infty I_\mM$ such that $\mN = P \mM_\infty P$ and
\begin{align}
\omega(G;\strategy') &= \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( Q^x_a \, Q^y_b \big)\notag\\
&=  \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a w\, w^* Q^y_b w\big)\notag\\
&\qquad\qquad+  \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big) \notag\\
&\qquad\qquad+ \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big(Q^x_a  Q^y_b (P-ww^*)\big)~.\label{eq:cv-0}
\end{align}
Here we used the fact that $Q^x_a \in \mN$ so $P Q^x_a P = Q^x_a$ for all $x,a$. 
For the second term on the right-hand side, we write
%\begin{align*}
%\Big|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y)  \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big)\Big|
%&=\Big| \tau\Big(  (P - w\, w^* )\Big( \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big)\Big)\Big|\\
%&\leq \tau(P-ww^*) \Big\|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big\|\\
%&\leq \eps\;,
%\end{align*}
%where the first inequality is by H\"older's inequality and the last inequality uses the definition of closeness to bound the first term. To bound the second term on the right-hand side, we write it as $\|A\circ B\|$, with $\circ$ the Hadamard product, and use  $\|A\circ B\| \leq \max_{i,j} |A_{i,j}|\|B\|$. Here $A = \sum_{a,b} D(a,b) P_{a,b}$, with $P_{a,b} = \sum_{i,j} \ket{u_{i,a}} \bra{v_{j,b}}$ with $\ket{u_{i,a}}$ a basis for the range of $Q^x_a$ and $\ket{v_{j,b}}$ for the range of $Q^y_b$, and $B=\sum_{a,b} Q^y_b ww^* Q^x_a = ww^*$.  
\begin{align}
\Big|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) & \tau\big( w^* Q^x_a (P - w\, w^* ) Q^y_b w\big)\Big|
\leq \Es{(x,y)\sim\mu} \Big | \tau\big( \sum_{a,b} D(a,b,x,y)   Q^y_b w w^* Q^x_a (P - w\, w^* ) \big) \Big | \;.\label{eq:XX1}
\end{align}
We then apply the Cauchy-Schwarz inequality
\begin{align}
\eqref{eq:XX1}&\leq \sqrt{ \tau((P - ww^*)^2)} \cdot \sqrt{\Es{(x,y)\sim\mu} \tau\big( \big| \sum_{a,b} D(a,b,x,y)   Q^x_a w w^* Q^y_b \big|^2 \big) } \notag\\
&\leq \sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu} \tau\big( \sum_{a,b} D(a,b,x,y)   Q^y_b ww^* Q^x_a w w^* Q^y_b \big) } \notag\\
&\leq \sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu}  \sum_{a,b}  \tau\big(Q^y_b ww^* Q^x_a w w^* Q^y_b \big) } \notag \\
&=\sqrt{\eps}\sqrt{ \Es{(x,y)\sim\mu}  \sum_{b}  \tau\big(Q^y_b ww^* \big) } \notag\\
&\leq \sqrt{\eps}\;.\label{eq:cv-1}
%&=\Big| \tau\Big(  (P - w\, w^* )\Big( \Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big)\Big)\Big|\\
%&\leq \tau(P-ww^*) \Big\|\Es{(x,y)\sim\mu} \sum_{a,b} D(a,b,x,y) Q^y_b ww^* Q^x_a \Big\|\\
%&\leq \eps\;,
\end{align}
The second line uses that $P - ww^*$ is a positive operator with operator norm at most $1$ for the first term, and that for fixed $x,y$, the measurements $\{ Q^x_a \}_a$ and $\{Q^y_b\}_b$ are projective. The third line is due to $D(x,y,a,b) \in \{0,1\}$. The fourt line is due to $\sum_a Q^x_a = I_\mN$. The fifth line is due to $\sum_a Q^y_b = I_\mN$. The third term on the right-hand side of~\eqref{eq:cv-0} is bounded in the same manner.

@@

In the next section we detail an important application of this test, Proposition~\ref{prop:dim-test}, which is used in the work~\cite{ji2020mip}. The present proof somewhat simplifies the one given in~\cite{ji2020mip}, but it keeps the same overall strategy. Moreover, Corollary~\ref{cor:qld} below is qualitatively stronger than~\cite[Theorem 7.14]{ji2020mip} as the latter applies to a slightly different game which in particular has a pair of questions with long answers, of $O(2^m\log q)$ bits, whereas the tester in Figure~\ref{fig:bRM-tester}, which is analyzed in the corollary, has short answers, $O(d)$ bits long; with a typical parameter setting (see the discussion following Theorem~\ref{thm:z2-stab}) this is exponentially smaller. The presence of long answers makes the analysis easier.

@@

To that end, assume that every genuine representation of $\Gamma$ is $(\delta,\mu_S)$-far from $\rho$. Namely, for every genuine representation $\varphi\colon \Gamma\to \mU(\mN)$, where $\mN=P\mM_\infty P$, and every isometry $w\in P\mM_\infty \Id_\mM$, we have 
\[
\max\left\{\Expectation_{s\sim \mu_S} \|\rho(s)-w^*\varphi(s)w\|_\tau^2\ ,\ \tau^\mM(\Id_\mM-w^*w)\right\}\geq \delta.
\]
Let $\varphi'\colon \Gamma\to \mU(\mN)$ be a genuine representation of $\Gamma$ which is $(\delta',\mu_S)$-close to $\rho'$, i.e.\ it satisfies 
\[
\frac{1}{|\Phi|}\Expectation_{s\sim \mu_S} \|\rho'(s)-w^*\varphi'(s)w\|_\tau^2\ ,\ \tau^{\mM_\Phi}(\Id_{\mM_\Phi}-w^*w)\leq  \delta',
\]
where  again  $\|X\|_\tau^2=\tau_\infty(X^*X)$.
As before, we can assume $w=PI_{\mM_\Phi}$. Then,
\[
\begin{split}
    \Expectation_{\alpha\in \Phi}\Expectation_{s\sim \mu_S} \|\rho(\alpha(s))-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2&=\frac{1}{|\Phi|}\Expectation_{s\sim \mu_S}\sum_{\alpha\in \Phi}\|\Id_{\mM_\alpha}\rho'(s)\Id_{\mM_\alpha}-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2 \\
    &\leq \frac{1}{|\Phi|}\Expectation_{s\sim \mu_S}\|\rho'(s)-\Id_{\mM_\Phi}\varphi'(s)\Id_{\mM_\Phi}\|_\tau^2\leq \delta'.
\end{split}
\]
In particular, by Markov's inequality, for at least two thirds of the $\alpha\in \Phi$ we have 
\[
\Expectation_{s\sim \mu_S} \|\rho(\alpha(s))-\Id_{\mM_\alpha}\varphi'(s)\Id_{\mM_\alpha}\|_\tau^2\leq 3\delta'.
\]
Similarly,
\[
\begin{split}
    \delta'&\geq \tau^{\mM_\Phi}(\Id_{\mM_\Phi}-\Id_{\mM_\Phi}P\Id_{\mM_\Phi})\\
    &=\Expectation_{\alpha\in\Phi}\tau_\infty(\Id_{\mM_\alpha}-\Id_{\mM_\alpha}P\Id_{\mM_\alpha})
\end{split}\;,
\]
and for at least two thirds of the $\alpha\in \Phi$ we have 
\[
\tau^{\mM_\alpha}(\Id_{\mM_\alpha}-\Id_{\mM_\alpha}P\Id_{\mM_\alpha})\leq 3\delta'.
\]
Hence, $\delta'\geq \frac{\delta}{3}$, and we deduce that every genuine representation of $\Gamma\cong \langle S\colon R\rangle$ is at least $(\nicefrac{\delta}{3},\infty)$-away from $\rho'$.
%Since the $L^\infty$ distance between functions is larger than the $L^1$ distance between them (with respect to any fully supported measure $\mu_S$), we conclude that a lower bound of $f(\eps)$ on the modulus of stability of a sufficiently symmetric presentation induces a lower bound of $\frac{\min{w_i}}{3}f(\eps)$ 
\begin{corollary}
    The transformation $\rho\mapsto\rho'$ we described  translates a $(\eps,\mu_R)$-approximate representation into a $((\max{\nicefrac{1}{w_i}})\cdot\eps,\infty)$-approximate representaion, and if every genuine representation of $\Gamma$ is $(\delta,\mu_S)$-far from $\rho$, then every genuine representation is $(\nicefrac{\delta}{3},\infty)$-far from $\rho'$.
\end{corollary}
\begin{remark}
    By applying this corollary to the construction from the beginning of the appendix, we conclude that the $L^\infty$ modulus of stability of the presentation \eqref{eq:z2-efficient} is $\Omega(k\eps)$.
\end{remark}

%%% AUTHOR:
%%% Bibliography goes here. Note that the arXiv cannot process bibtex
%%% or biber bibliographies.  Example of acceptable bibliograpy format:

\bibliography{qld}

%
%\bibliographystyle{amsplain}
%\begin{thebibliography}{99}
%\bibitem{bergelson-johnson-moreira}
%Vitaly Bergelson, John H. Johnson Jr., and Joel Moreira.
%\newblock New polynomial and multidimensional extensions of classical partition
  %results.
%\newblock 2015, arXiv:1501.02408.
%
%\bibitem{cilleruelo}
%Javier Cilleruelo.
%\newblock Combinatorial problems in finite fields and {S}idon sets.
%\newblock {\em Combinatorica}, 32(5):497--511, 2012.
%
%\end{thebibliography}

%% AUTHOR: You can generate such a bibliography from a .bib file by 
%% running pdflatex/bibtex/pdflatex/pdflatex and then pasting the .bbl file
%% between \begin{thebibliography} and \end{bibliography}

%%% AUTHOR: Include a short description of each author following the
%%% structure below. Use the same short tags used previously.  
%%% Use \imageat{} and \imagedot{} instead of "@" and "." in
%%% email addresses-this replaces the symbols with graphics to avoid 
%%% e-mail address harvesting from the .pdf file
\begin{dajauthors}
\begin{authorinfo}[mc]
  Michael Chapman\\
	Member of the school of Mathematics\\
  Institute for Advanced Study\\
  Princeton, New Jersey, USA\\
  mchapman\imageat{}ias\imagedot{}edu \\
\end{authorinfo}
\begin{authorinfo}[tv]
  Thomas Vidick\\
  Professor\\
  École Polytechnique Fédérale de Lausanne\\
  Lausanne, Switzerland\\
  thomas\imagedot{}vidick\imageat{}epfl\imagedot{}ch \\
\end{authorinfo}
\begin{authorinfo}[hy]
  Henry Yuen\\
  Associate Professor\\
   Columbia University\\
  New York, New York, USA\\
  hyuen\imageat{}cs\imagedot{}columbia\imagedot{}edu\\
\end{authorinfo}
\end{dajauthors}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "daj-template"
%%% End:
